The Adaptive Multi-scale Simulation Infrastructure - Design and Implementation of a soft tissue simulation

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
removed load balancing section

\subsection{Soft Tissue Simulation}\label{biotissue}

The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation. The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the global simulation. Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh, see figure \ref{biotissue_hierarchy}.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
  \end{center}

  \caption{\small biotissue multi-scale problem}
  \label{biotissue_hierarchy}
\end{figure}

@@

Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.

\subsection{Biotissue Scale-Sensitive Load Balancing}\label{biotissue_load_balancing}

Scale-sensitive load balancing is conducted on the micro-scale RVE distribution at locations in the code where communication bottlenecks are already present. This prevents the introduction of new bottlenecks which could result in reduced performance. Macro-scale Newton-iterations are the most frequently occuring multi-scale communication bottleneck at which we chose to conduct the load balancing operation, and macro-scale incremental load steps the least-frequent. These represent the two extreme cases considered in our analysis on the efficacy of the scale-sensitive load-balancing scheme adopted for the biotissue simulation.

Each micro-scale RVE simulation involves the assembly and solve of a linear system of equations on the order of 10,000 unknowns for each micro-scale Newton iteration. While the total solve time is dependent on the number of nonlinear iterations and the hardware executing the code, this system is sufficiently small that parallelization of the RVE code itself is unwarranted, as the introduction of communication overhead into the RVE solution process would result in decreased performance, despite the modest gain to computation time. However, as each RVE represents an independent task to be accomplished by the simulation, they may be freely distributed across the processes assigned to the micro-scale computation. The only restriction to their (re)distribution for load balancing purposes is that the relation to a specific macro-scale numerical integration point must be maintained.

As each RVE is a nonlinear simulation their is no a-prior estimate of the computational load each RVE represents, so the optimal initial distribution is simply treating the weight of each RVE as 1. Thus each micro-scale process is assigned either $\floor*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$ or $\ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$. Once the first load-balancing event is reached, regardless of the granularity at which load-balancing is occuring for a given case, each RVE has recorded the total number of newton iterations it has conducted since the previous load-balancing event. This number is used as a heuristic weighting metric to determine the computational load of each individual RVE. The iteration count metric was chosen as each iteration takes a similar level of work to complete, up to the difference in the number of RVEs for each unique fiber network. Further, assuming linear incremental loading, regions in the engineering-scale mesh located near stress concentrators will result in 'more interesting' boundary conditions for the micro-scale problems related to numerical integration points in that region, which may result in the RVE requiring more Newton iterations to converge to an adequate solution, since the boundary conditions may lie farther from the initial state of the RVE in the regime of convergence for the nonlinear problem.

\section{Results}\label{results}

>>>>

\subsection{Soft Tissue Simulation}\label{biotissue}
The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation. The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the global simulation. Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
  \end{center}
  \caption{\small biotissue multi-scale problem}
  \label{biotissue_hierarchy}
\end{figure}

@@

Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process. 
\section{Results}\label{results}
